Symmetry-based constraints on the size of the dynamical Solar neighborhood

### 2019-04-09 paper/ms.tex
Minor changes throughout

\shortauthors{Beane et al.}
\shorttitle{The Midplane is Not a Plane}

@@

\title{The Midplane is Not a Plane: Implications for {\em Gaia}}

@@

If the system being considered departs from axisymmetry in a significant and non-adiabatic way, the actions computed using an axisymmetric approximation to the true potential can exhibit cyclic dependence on the orbital phase (or time at which the star's position and velocity are observed), large-scale migration, or diffusion from their initial values.
In the Milky Way, stellar actions are expected to diffuse on short time scales due to scattering with gas clouds
and to evolve on longer time scales in the case of orbits near resonances with spiral arms, bar(s), and other
large scale perturbations \citep{2014RvMP...86....1S}. For this reason, we and
other authors have used actions to study stellar scattering in the Milky Way
disk using the improved astrometry of {\em Gaia} DR2 and various age
catalogues \citep{2018ApJ...867...31B,2018arXiv180803278T}. Actions have also
been used to study different models of spiral structure in the Milky Way
\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in the extended Solar neighborhood in action space are discussed in
\citet{2018arXiv180503653T}.

@@

In this work we discuss how the imposition of a global axisymmetric coordinate system based on a local measurement of the midplane location and Galactocentric distance introduces systematic error in the computation of actions, especially for stars in the thin disk. Our main point
is that the ``local'' Galactic midplane is not the same as the ``global''
Galactic midplane. For stars relatively far from the Sun,
the local midplane location will not agree with
our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution, not necessarily even as large as spiral arms. Converting positions and velocities of more distant stars from a heliocentric to a
Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate relative to $z$ velocity, and hence the actions computed for these stars. The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$. Depending on the type of orbit being studied, this sets a limit on the ``dynamical solar neighborhood'' over which the local coordinate frame can be usefully extended for action-space analysis. A similar argument applies to uncertainty in measurements of
$R_{\odot}$ and variations in the local standard of rest (LSR). In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions.

@@

In Section~\ref{sec:discussion} we discuss the implications of midplane variations, and the resulting systematic uncertainty in the vertical action, for action-space analyses beyond the dynamical solar neighborhood. As an illustration, in Section \ref{sssec:neighborhood} we use these estimates to consider the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions. In light of our findings, we considered the orbits of the
nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
collaboration \citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters}, we
report the distance to each cluster, the three actions, and the maximum
vertical extent of their orbits ($z_{\text{max}}$). 
%The actions are computed using the technique described in Section~\ref{ssec:action_comp}. 
 In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if background is unimportant.  We summarize and conclude in Section~\ref{sec:conclusion}.

@@

Our main point is that the local midplane varies between different points in
the Galaxy. This is a different statement than that our measurements of the
local midplane are erroneous. Indeed, the local midplane has recently been
measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
that determinations of the local midplane are perfect, since midplane
measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
additional complication is the fact that non-zero Solar heights induce a tilt
in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}. We do not propose here what a modeler should use for the ``true'' global midplane, since it depends on the particular problem being studied. Even without a high-precision definition of the global Galactic midplane, current observations from Gaia should soon permit a measurement of the real azimuthal dependence of the midplane location.

>>>>

\shortauthors{Beane et al.}
\shorttitle{The dynamical solar neighborhood}

@@

\title{Symmetry-based constraints on the size of the dynamical Solar neighborhood}

@@

If the system being considered departs from axisymmetry in a significant and/or non-adiabatic way, the actions computed using an axisymmetric approximation to the true potential can exhibit cyclic dependence on the orbital phase (or time at which the star's position and velocity are observed), large-scale migration, or diffusion from their initial values.
In the Milky Way, stellar actions are expected to diffuse on short time scales due to scattering with gas clouds
and to evolve on longer time scales in the case of orbits near resonances with spiral arms, bar(s), and other
large scale perturbations \citep{2014RvMP...86....1S}. For this reason, we and
other authors have used actions to study stellar scattering in the Milky Way
disk using the improved astrometry of {\em Gaia} DR2 and various age
catalogues \citep{2018ApJ...867...31B,2018arXiv180803278T}. Actions have also
been used to study different models of spiral structure in the Milky Way
\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in the extended Solar neighborhood in action space are discussed in
\citet{2018arXiv180503653T}.

@@

In this work we discuss how the imposition of a global axisymmetric coordinate system based on a local measurement of the midplane location and Galactocentric distance introduces systematic error in the computation of actions, especially for stars in the thin disk. Our main point
is that the ``local'' Galactic midplane is not the same as the ``global''
Galactic midplane. For stars relatively far from the Sun,
the local midplane location will not agree with
our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution. Converting positions and velocities of more distant stars from a heliocentric to a
Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate, and hence the actions computed for these stars. The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$. Depending on the type of orbit being studied, this sets a limit on the ``dynamical solar neighborhood'' over which the local coordinate frame can be usefully extended for action-space analysis. A similar argument applies to uncertainty in measurements of
$R_{\odot}$ and variations in the local standard of rest (LSR). In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions.

@@

In Section~\ref{sec:discussion} we discuss the implications of midplane variations, and the resulting systematic uncertainty in the vertical action, for action-space analyses beyond the dynamical solar neighborhood. As an illustration, in Section \ref{sssec:neighborhood} we use these estimates to consider the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions. In light of our findings, we considered the orbits of the
nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
collaboration \citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters}, we
report the distance to each cluster, the three actions, and the maximum
vertical extent of their orbits ($z_{\text{max}}$). 
%The actions are computed using the technique described in Section~\ref{ssec:action_comp}. 
 In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if contamination from foreground or background stars is unimportant. We summarize and conclude in Section~\ref{sec:conclusion}.

@@

Our main point is that the local midplane varies between different points in
the Galaxy. This is a different statement than that our measurements of the
local midplane are erroneous. Indeed, the local midplane has recently been
measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
that determinations of the local midplane are perfect, since midplane
measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
additional complication is the fact that non-zero Solar heights induce a tilt
in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}. We do not propose here what a modeler should use for the ``true'' global midplane, since it depends on the particular problem being studied. Even without a high-precision definition of the global Galactic midplane, current observations from {\em Gaia} should soon permit a measurement of the real azimuthal dependence of the midplane location.
